---
title: "RSTR_LuminexAnalysis_WMtbL"
author: "Nathan Kieswetter"
date: "2024-05-16"
output: html_document
---

## Load in the required libraries

```{r}
library(ComplexHeatmap)
library(tidyverse)
library(ggpubr)
library(gridExtra)
library(circlize)
```

## Read in the counts

```{r}
df_RSTR_comb <- read.csv("G:\\data\\here\\RSTR_Luminex\\out\\luminex\\RSTR_luminex_all_batches.csv")
```

## Log-transform mfi data, run paired-T-test, extract the paired T-statistic, and plot a heatmap

```{r}
df_RSTR_comb_test <- df_RSTR_comb %>%
  mutate_at(vars(final_MFI, DMSO_final_MFI), list(~ifelse(. < 0, 0, .))) %>%
  mutate(final_MFI_log = log1p(final_MFI),
         DMSO_final_MFI_log = log1p(DMSO_final_MFI)
  )     
```

Log10 transformed background connected MFI

```{r}
df_RSTR_comb_test <- df_RSTR_comb_test %>%
  mutate(backcorr_MFI_log = final_MFI_log - DMSO_final_MFI_log)
```

Split the data into subgroups to run the paired-sample T-test

```{r}
df_list <- df_RSTR_comb_test %>%
  split(f = list(df_RSTR_comb_test$stimulation,
                 df_RSTR_comb_test$Analyte,
                 df_RSTR_comb_test$timepoint))

T.test_results <- lapply(df_list, function(x) with(x, t.test(x = final_MFI_log, 
                                                             y = DMSO_final_MFI_log, 
                                                             paired = TRUE, 
                                                             alternative = "two.sided",
                                                             na.rm = TRUE)))
```

Extract the T-test results and create a data frame of these data

```{r}
names_list <- names(T.test_results)

df_ttest_results <- T.test_results %>% 
  map_df(broom::tidy) %>%
  cbind(names_list) %>%
  separate(names_list, c('stimulation', 'analyte', 'timepoint'),"[.]") %>%
  select(stimulation, analyte, timepoint, statistic, p.value, method) %>%
  drop_na() %>%
  mutate(combined = str_c(stimulation, timepoint, sep = ".", collapse = NULL)) %>%
  filter(stimulation == "MTBL") 

df_list_2 <- df_ttest_results %>%
  split(f = list(df_ttest_results$stimulation,
                 df_ttest_results$analyte))

df_mean_list <- lapply(df_list_2, function(df) {
  df_mean <- df %>%
    group_by(timepoint) %>%
    summarize(
      statistic_mean = mean(statistic),
      id = first(combined),
      analyte = first(analyte))
}) %>%
  bind_rows() %>%
  separate(id, c("stimulation"), sep = "\\.") %>%
  filter(stimulation == "MTBL") # Only select the MTBL stimulated samples for the RSTR sequel paper

df_ttest_mat<- df_mean_list %>%
  pivot_wider(names_from = analyte, values_from = statistic_mean) %>%
  unite(col = "combined", timepoint, stimulation, sep = ".") %>%
  column_to_rownames("combined") 
```

## Portrait of heatmap matrix for publication

```{r}
df_ttest_mat_port <- t(df_ttest_mat)

df_ttest_mat_port <- df_ttest_mat_port %>%
  as.data.frame()%>%
  select("6.MTBL", "12.MTBL", "24.MTBL", "48.MTBL" )
```

## Extract the metadata for use in the heatmap

```{r}
metadata <- df_ttest_mat %>%
  as.data.frame() %>%
  rownames_to_column() %>%
  separate(rowname, c('timepoint', 'stimulation'),"[.]") %>%
  select(timepoint)

# Set the annotations

ann <- metadata %>%
  data.frame()

#colnames(ann) <- c("stimulation", "timepoint")

colAnn <- rowAnnotation(df = ann,
                        annotation_width = unit(c(1, 4), 'cm'),
                        gp = gpar(col = "white", lwd = 2),
                        gap = unit(1, 'mm'),
                        col = list('timepoint' = c('6' = 'lightblue4', 
                                                   '12' = 'lightgoldenrod', 
                                                   '24' = 'lightgray', 
                                                   '48' = 'lightpink3')
                        ))

col_fun = colorRamp2(c(-10, 0, 15), c("blue", "white", "red"))
lgd = Legend(col_fun = col_fun, title = "Paired T-statistic")

Heatmap(df_ttest_mat,
        # cluster_rows = group,
        name = "Paired T-statistic", 
        right_annotation = colAnn,
        show_column_dend = TRUE,
        show_row_dend = TRUE,
        show_row_names = FALSE,
        rect_gp = gpar(col = "white", lwd = 2),
        cell_fun = function(j, i, x, y, width, height, fill) {
          grid.text(sprintf("%.1f", df_ttest_mat[i, j]), x, y, gp = gpar(fontsize = 10))
        })
```

## Portrait Heat Map

```{r}
hm <- Heatmap(df_ttest_mat_port, 
        # cluster_rows = group,
        name = "Paired T-statistic", 
#        right_annotation = colAnn,
        show_column_dend = FALSE,
        show_row_dend = FALSE,
        show_row_names = TRUE,
        column_order = order(as.numeric(gsub(".PP1", "", colnames(df_ttest_mat_port)))),
#         row_order = c("GM-CSF","IFNa",  "IFNy","IL-10", "IL-12p70","IL-13","IL-15","IL-17A","IL-18","IL-1a",
#                       "IL-1b","IL-1RA","IL-2","IL-21","IL-22","IL-23","IL-27","IL-31","IL-4","IL-5","IL-6",     
#                       "IL-7","IL-8","IL-9","IP-10","TNFa","TNFb"),  
        rect_gp = gpar(col = "white", lwd = 2),
        # cell_fun = function(j, i, x, y, width, height, fill) {
        #   grid.text(sprintf("%.1f", df_ttest_mat_port[i, j]), x, y, gp = gpar(fontsize = 10))
        # }
)

hm

row_order_hm <- row_order(hm)
```

## Extract the time points the highest signal-noise ratio

```{r}
df_mean_list_2 <- lapply(df_list_2, function(df) {
  df_mean <- df %>%
    group_by(timepoint) %>%
    summarize(
      statistic_mean = mean(statistic),
      id = first(combined),
      analyte = first(analyte))}) %>%
  bind_rows() %>%
  separate(id, c("stimulation"), sep = "\\.") %>%
  mutate(statistic_mean = scale(statistic_mean)) %>% #scaled to fit on 
  split(f = list(df_mean_list$stimulation,
                 df_mean_list$analyte))
```

Function to select the time point with the highest signal to noise ratio

```{r}
select_highest_row <- function(df, col_name) {
  df[which.max(df[[col_name]]),]
}

# Determine the best time points

highest_rows <- lapply(df_mean_list_2, 
                       select_highest_row, 
                       col_name = "statistic_mean") %>%
                bind_rows()
```

Select the MTBL stimulated samples and Subset the original data by the selected time points calculated in "highest rows"

```{r}
df_RSTR_comb_test_MTBL <- df_RSTR_comb_test %>%
  filter(stimulation == "MTBL") %>% #only interested in Mtb lysate
  mutate(group = recode(group,'P_neg' = 'RSTR',
                        'TST_pos' = 'LTBI'))

# Loop to pull out timepoints with best signal to noise ratio for plotting!
df_to_plot <- list()

# iterate through each row in highest_rows
for (i in 1:nrow(highest_rows)) {
  
  # get the current row data in highest_rows
  current_row <- highest_rows[i, c("timepoint", "analyte", "stimulation")]
  
  # find rows in df_RSTR_comb_test that match the current row based on timepoint, analyte, and stimulation
  matching_rows <- df_RSTR_comb_test_MTBL %>% 
    filter(timepoint == current_row$timepoint,
           Analyte == current_row$analyte,
           stimulation == current_row$stimulation)
  
  # add the matching rows to the matched_rows list
  if (nrow(matching_rows) > 0) {
    df_to_plot[[i]] <- matching_rows
  } else {
    df_to_plot[[i]] <- data.frame(matrix(nrow = 0, ncol = ncol(df_RSTR_comb_test_MTBL)))
  }
}

df_to_plot <- do.call(rbind, df_to_plot) # combine the matched rows into a single dataframe

# Splits df into multiple df's based on analyte (for plotting)

analyte_list <- split(df_to_plot, f = df_to_plot$Analyte)

tmp <- df_to_plot %>%
  filter(stimulation != "DMSO")
#analyte_list_noDMSO <-  split(tmp, f = tmp$Analyte)

analyte_list_noDMSO <- split(tmp, f = list(tmp$Analyte))
```

## Plot all cytokines as a measure background corrected MFI values over time (i.e. DMSO subtracted)

```{r}
plotdata3 <- function(x) { 
  ggplot(analyte_list_noDMSO[[x]], 
         aes(x = factor(group, levels = c("RSTR", "LTBI")), 
             y = backcorr_MFI_log,
             color = group)) + 
    ggtitle(names(analyte_list_noDMSO)[[x]]) +
 #   labs(fill = 'Group') +
    geom_boxplot(color = "black", outlier.shape = NA) +
    geom_point(pch = 1, size = 2, lwd = 50, 
               position = position_jitterdodge(jitter.width = 0.3)) +
    xlab(NULL) + #comment this out as needed
    ylab(NULL) + #comment this out as needed
    # xlab("Group") +                             #comment this out for grid
    # ylab("Background Corrected MFI (Log10)") + # comment this out for grid
    theme(plot.title = element_text(hjust = 0.5))+
    theme_bw() +
    theme(text = element_text(family="Arial"),
          # axis.text.x = element_text(size = 15),  
          plot.title = element_text(hjust = 0.5),
          legend.position = "none",
          axis.text.x = element_blank(), #comment this out as needed
          axis.ticks.x = element_blank() #comment this out to remove axis
    ) + 
    ylim(NA, max(analyte_list_noDMSO[[x]]$backcorr_MFI_log) + 0.5) +
    xlab(NULL) + #comment this out as needed
    ylab(NULL) + #comment this out as needed
    scale_color_manual(values = c("#4DAF4A","#984EA3")) +
    stat_compare_means(method = "t.test",
                       paired = FALSE,
                       label = "p.signif",
                       label.x = 1.5,
                       label.y = max(analyte_list_noDMSO[[x]]$backcorr_MFI_log) - 0.01, 
                       symnum.args = list(cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 1),
                                          symbols = c("****", "***", "**", "*", "ns"))
)
}

List_2 <- lapply(seq_len(length(analyte_list_noDMSO)), plotdata3)
```

## Arrange graphs in a grid to match the order in the clustered heatmap

```{r}
highest_rows2 <- highest_rows %>%
  mutate(row_num = row_number()) %>% 
  mutate(row_num = factor(row_num, levels = as.character(as.numeric(row_order_hm)))) %>% 
  arrange(row_num) %>%
  column_to_rownames(var = "row_num")


# Define the order of the plots

plot_order <- as.numeric(row.names(highest_rows2))

# Rearrange the list of plots

new_list <- List_2[plot_order]

# Generate the arranged plots using grid.arrange()

grid.arrange(grobs = matrix(new_list),
             ncol = 5,
             left = gridtext::richtext_grob("Background Subtracted MFI (log10)", rot = 90),
             bottom = gridtext::richtext_grob(text =  ' Groups:  <span style="color:purple"> RSTR</span>  <span style="color:limegreen">LTBI</span>'),
             clip = FALSE
             )
```

## Export all the individual SVG files

IL-23 at 6 hours for publication

```{r}
df_to_plot_mtbl_IL23 <- df_RSTR_comb_test %>%
  filter(stimulation == "MTBL",
         Analyte == "IL-23",
         timepoint == "6",
         stimulation != "DMSO") %>%
  mutate(group = recode(group,'P_neg' = 'RSTR',
                        'TST_pos' = 'LTBI'))

ggplot(df_to_plot_mtbl_IL23, 
       aes(x = factor(group, levels = c("RSTR", "LTBI")), 
           y = backcorr_MFI_log,
           color = group)) + 
  ggtitle("IL-23") +
  #   labs(fill = 'Group') +
  geom_boxplot(color = "black", outlier.shape = NA) +
  geom_point(pch = 1, size = 2, lwd = 50, 
             position = position_jitterdodge(jitter.width = 0.3)) +
  xlab("Group") +                             #comment this out for grid
  ylab("Background Corrected MFI (Log10)") + # comment this out for grid
  theme(plot.title = element_text(hjust = 0.5))+
  theme_bw() +
  theme(text = element_text(family="Arial"),
        # axis.text.x = element_text(size = 15),  
        plot.title = element_text(hjust = 0.5),
        legend.position = "none",
        # axis.text.x = element_blank(), #comment this out as needed
        # axis.ticks.x = element_blank() #comment this out to remove axis
  ) + 
  scale_color_manual(values = c("#4DAF4A","#984EA3")) +
  stat_compare_means(method = "t.test",
                     paired = FALSE,
                     label.x = 1.3,
                     # label = "p.signif",
                     # symnum.args = list(cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 1),
                     #                    symbols = c("****", "***", "**", "*", "ns"))
  )
```

IL-17a at 24 hours

```{g}
df_to_plot_mtbl_IL17 <- df_RSTR_comb_test %>%
  filter(stimulation == "MTBL",
         Analyte == "IL-17A",
         timepoint == "24",
         stimulation != "DMSO") %>%
  mutate(group = recode(group,'P_neg' = 'RSTR',
                        'TST_pos' = 'LTBI'))

ggplot(df_to_plot_mtbl_IL17, 
       aes(x = factor(group, levels = c("RSTR", "LTBI")), 
           y = backcorr_MFI_log,
           color = group)) + 
  ggtitle("IL-17A") +
  #   labs(fill = 'Group') +
  geom_boxplot(color = "black", outlier.shape = NA) +
  geom_point(pch = 1, size = 2, lwd = 50, 
             position = position_jitterdodge(jitter.width = 0.3)) +
  xlab("Group") +                             #comment this out for grid
  ylab("Background Corrected MFI (Log10)") + # comment this out for grid
  theme(plot.title = element_text(hjust = 0.5))+
  theme_bw() +
  theme(text = element_text(family="Arial"),
        # axis.text.x = element_text(size = 15),  
        plot.title = element_text(hjust = 0.5),
        legend.position = "none",
        # axis.text.x = element_blank(), #comment this out as needed
        # axis.ticks.x = element_blank() #comment this out to remove axis
  ) + 
  scale_color_manual(values = c("#4DAF4A","#984EA3")) +
  stat_compare_means(method = "t.test",
                     paired = FALSE,
                     label.x = 1.3,
                     # label = "p.signif",
                     # symnum.args = list(cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 1),
                     #                    symbols = c("****", "***", "**", "*", "ns"))
  )
```

IL-10 at 12 hours

```{r}
df_to_plot_mtbl_IL10 <- df_RSTR_comb_test %>%
  filter(stimulation == "MTBL",
         Analyte == "IL-10",
         timepoint == "12",
         stimulation != "DMSO") %>%
  mutate(group = recode(group,'P_neg' = 'RSTR',
                        'TST_pos' = 'LTBI'))

ggplot(df_to_plot_mtbl_IL10, 
       aes(x = factor(group, levels = c("RSTR", "LTBI")), 
           y = backcorr_MFI_log,
           color = group)) + 
  ggtitle("IL-10") +
  #   labs(fill = 'Group') +
  geom_boxplot(color = "black", outlier.shape = NA) +
  geom_point(pch = 1, size = 2, lwd = 50, 
             position = position_jitterdodge(jitter.width = 0.3)) +
  xlab("Group") +                             #comment this out for grid
  ylab("Background Corrected MFI (Log10)") + # comment this out for grid
  theme(plot.title = element_text(hjust = 0.5))+
  theme_bw() +
  theme(text = element_text(family="Arial"),
        # axis.text.x = element_text(size = 15),  
        plot.title = element_text(hjust = 0.5),
        legend.position = "none",
        # axis.text.x = element_blank(), #comment this out as needed
        # axis.ticks.x = element_blank() #comment this out to remove axis
  ) + 
  scale_color_manual(values = c("#4DAF4A","#984EA3")) +
  stat_compare_means(method = "t.test",
                     paired = FALSE,
                     label.x = 1.3,
                     # label = "p.signif",
                     # symnum.args = list(cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 1),
                     #                    symbols = c("****", "***", "**", "*", "ns"))
                     
  )
```
