---
title: "RSTR_TGFb_ELISA_WMtbL"
author: "Nathan Kieswetter_Seshadri Lab"
date: "2024-05-20"
output: html_document
---

Load in appropriate libraries

```{r setup, include=FALSE}
library(ComplexHeatmap)
library(tidyverse)
library(ggpubr)
library(svglite)
```

# Read in the TGFb analysis count data

```{r}

df_RSTR_comb <- read.csv("G:\\Shared drives\\Seshadri Lab\\Lab Members\\Kieswetter_Nathan\\data\\here\\RSTR_Luminex\\data\\elisa\\RSTR_TGFb_all_batches_mfi.csv")
```

# Log-transform mfi data, run paired-T-test, extract the paired T-statistic, and plot a heatmap 

```{r}
df_RSTR_comb <- df_RSTR_comb %>%
  mutate_at(vars(mfi_final, DMSO_final_MFI), list(~ifelse(. < 0, 0, .))) %>%
  mutate(final_MFI_log = log1p(mfi_final),
         DMSO_final_MFI_log = log1p(DMSO_final_MFI)
  )                         
```

# Add a column for the Log10 transformed background connected MFI

```{r}
df_RSTR_comb <- df_RSTR_comb %>%
  mutate(backcorr_MFI_log = final_MFI_log - DMSO_final_MFI_log)
```

# Split the data into subgroups to run the paired-sample T-test

```{r}
df_list <- df_RSTR_comb %>%
  split(f = list(df_RSTR_comb$stimulation,
                 df_RSTR_comb$timepoint))

T.test_results <- lapply(df_list, function(x) with(x, t.test(x = final_MFI_log, 
                                                             y = DMSO_final_MFI_log, 
                                                             paired = TRUE, 
                                                             alternative = "two.sided",
                                                             na.rm = TRUE)))
```

# Extract the T-test results and create a data frame of these data

```{r}
names_list <- names(T.test_results)

df_ttest_results <- T.test_results %>% 
  map_df(broom::tidy) %>%
  cbind(names_list) %>%
  separate(names_list, c('stimulation', 'timepoint'),"[.]") %>%
  select(stimulation, timepoint, statistic, p.value, method) %>%
  drop_na() %>%
  mutate(combined = str_c(stimulation, timepoint, sep = ".", collapse = NULL)) %>%
  filter(stimulation != "DMSO", #%>% # Remove any DMSO comparisons
         stimulation == "MTBL") 

df_list_2 <- df_ttest_results %>%
  split(f = list(df_ttest_results$stimulation))

df_mean_list <- lapply(df_list_2, function(df) {
  df_mean <- df %>%
    group_by(timepoint) %>%
    summarize(
      statistic_mean = mean(statistic),
      id = first(combined))
}) %>%
  bind_rows() %>%
  separate(id, c("stimulation"), sep = "\\.") %>%
  filter(stimulation == "MTBL") %>%
  mutate(analyte = "TGFb")# Only select the PP1 stimulated samples for the RSTR sequel paper

df_ttest_mat<- df_mean_list %>%
  pivot_wider(names_from = analyte, values_from = statistic_mean) %>%
  unite(col = "combined", timepoint, stimulation, sep = ".") %>%
  column_to_rownames("combined") 
```

# Extract the metadata for use in the heat map

```{r}
metadata <- df_ttest_mat %>%
  as.data.frame() %>%
  rownames_to_column() %>%
  separate(rowname, c('timepoint', 'stimulation'),"[.]") %>%
  select(timepoint)

# Set the annotations

ann <- metadata %>%
  data.frame()

colAnn <- rowAnnotation(df = ann,
                        annotation_width = unit(c(1, 4), 'cm'),
                        gp = gpar(col = "white", lwd = 2),
                        gap = unit(1, 'mm'),
                        col = list('timepoint' = c('6' = 'royalblue', 
                                                   '12' = 'blueviolet', 
                                                   '24' = 'darkslateblue', 
                                                   '48' = 'navyblue')
                        ))
my_row_order <- c("24.PP1", "48.PP1", "12.PP1", "6.PP1")

Heatmap(df_ttest_mat, 
        # cluster_rows = group,
        name = "Paired T-statistic", 
        right_annotation = colAnn,
        row_order = my_row_order,    # this works, but is generally non-informative
        show_column_dend = FALSE,
        show_row_dend = FALSE,
        cluster_rows = FALSE, 
        cluster_columns = FALSE,
        show_row_names = FALSE,
        rect_gp = gpar(col = "white", lwd = 2),
        cell_fun = function(j, i, x, y, width, height, fill) {
          grid.text(sprintf("%.1f", df_ttest_mat[i, j]), x, y, gp = gpar(fontsize = 10))
        })
```

# Extract the time points the highest signal-noise ratio

```{r}
df_to_plot_mtbl <- df_RSTR_comb %>%
  filter(stimulation == "MTBL",
         timepoint == "6",
         stimulation != "DMSO") %>%
  mutate(group = recode(group,'P_neg' = 'RSTR',
                        'TST_pos' = 'LTBI'))

### MTBL
ggplot(df_to_plot_mtbl, 
       aes(x = factor(group, levels = c("RSTR", "LTBI")), 
           y = backcorr_MFI_log,
           color = group)) + 
  labs(fill = 'Group') +
  geom_boxplot(color = "black", outlier.shape = NA) +
  geom_point(pch = 1, size = 2, lwd = 50, 
             position = position_jitterdodge(jitter.width = 0.3)) +
  ggtitle("TGFb*") +
  xlab(NULL) +
  ylab(NULL) +
  # xlab("Group") +
  # ylab("Background Corrected MFI (Log10)") +
  theme(plot.title = element_text(hjust = 0.5),)+
  theme_bw() +
  theme(text = element_text(family="Arial"),
        plot.title = element_text(hjust = 0.5),
        legend.position = "none",
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()
  ) + 
  scale_color_manual(values = c("#4DAF4A","#984EA3")) +
  stat_compare_means(method = "wilcox.test",
                     label = "p.signif",
                     label.x = 1.4,
                     symnum.args = list(cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 1),
                                        symbols = c("****", "***", "**", "*", "ns")))
```
